---
title: "Introduction to Functions"
output:
  html_document:
    theme: readable
    highlight: tango
    self_contained: false
---


A function is a basic data recipe. It is also an input-output machine. It accepts arguments, and returns the final product of the recipe.

Let's create a mortgage calculator as a simple example. 

$$
monthly \ payments = \frac{principle \cdot interest \ rate}{1-(1+interest \ rate)^{-loan \ duration}}
$$

The function will look like this:

```{r eval=FALSE}
 calcMortgage <- function( principal, years, APR )
 {
    months <- years * 12
    int.rate <- APR / 12
    
    # amortization formula
    monthly.payment <- ( principal * int.rate ) / 
                       (1 - (1 + int.rate)^(-months) )
    
    monthly.payment <- round( monthly.payment, 2 )
    
    return( monthly.payment )
 
 }
```



```{r, include=FALSE}
tutorial::go_interactive( greedy=FALSE )
```

```{r ex="example-01", type="pre-exercise-code", tut=TRUE}
 calcMortgage <- function( principal, years, APR )
 {
    months <- years * 12
    int.rate <- APR / 12
    
    # amortization formula
    monthly.payment <- ( principal * int.rate ) / 
                       (1 - (1 + int.rate)^(-months) )
    
    monthly.payment <- round( monthly.payment, 2 )
    
    return( monthly.payment )
 
 }
```




```{r ex="example-01", type="sample-code", tut=TRUE}
# Check the function's arguments:
args( calcMortgage )

# Calculate a mortgage for a $100,000 house for 30 years with a 4.5 percent interest rate (0.045 APR)
calcMortgage(  )  # add arguments here

# Try: calcMortgage( principal=100000 )
calcMortgage( principal=100000 )
```


What happens if you omit an argument from the function?

Why is that? 




# Default Arguments

We can add default values for arguments. These defaults allow us to utilize the function without specifying values for those arguments.

```{r eval=FALSE}
 calcMortgage <- function( principal, years=30, APR=0.045 )
 {
    months <- years * 12
    int.rate <- APR / 12
    
    # amortization formula
    monthly.payment <- ( principal * int.rate ) / 
                       (1 - (1 + int.rate)^(-months) )
    
    monthly.payment <- round( monthly.payment, 2 )
    
    return( monthly.payment )
 
 }
```


```{r ex="example-02", type="pre-exercise-code", tut=TRUE}
 calcMortgage <- function( principal, years=30, APR=0.045 )
 {
    months <- years * 12
    int.rate <- APR / 12
    
    # amortization formula
    monthly.payment <- ( principal * int.rate ) / 
                       (1 - (1 + int.rate)^(-months) )
    
    monthly.payment <- round( monthly.payment, 2 )
    
    return( monthly.payment )
 
 }
```




```{r ex="example-02", type="sample-code", tut=TRUE}
# Notice the new default arguments:

args( calcMortgage )


# Calculate a mortgage for a $100,000 house for 30 years with a 4.5 percent APR

calcMortgage( )

# Now try: calcMortgage( principal=100000 )

calcMortgage( principal=100000 )
```


Note that `calcMortgage( principal=100000 )` now works because the function uses the default values for years and APR. 

Can you still use custom values for those arguments after defaults are set? 



## Defaults

```{r ex="example-03", type="pre-exercise-code", tut=TRUE}
 calcMortgage <- function( principal, years=30, APR=0.045 )
 {
    months <- years * 12
    int.rate <- APR / 12
    
    # amortization formula
    monthly.payment <- ( principal * int.rate ) / 
                       (1 - (1 + int.rate)^(-months) )
    
    monthly.payment <- round( monthly.payment, 2 )
    
    return( monthly.payment )
 
 }
```

```{r ex="example-03", type="sample-code", tut=TRUE}
# Calculate a mortgage for a $100,000 house for 15 years with a 3 percent APR

calcMortgage()
```


# Implicit Argument Calls

An explicit call to arguments always uses the formal argument name such as `principal=100000`.

You can, however, also use implicit argument calls. These rely on the order arguments are specified in the function, and the order of your values in your call.

Implicit arguments calls can be risky, however, because it is very easy to mix them up. 


```{r ex="example-04", type="pre-exercise-code", tut=TRUE}
 calcMortgage <- function( principal, years=30, APR=0.045 )
 {
    months <- years * 12
    int.rate <- APR / 12
    
    # amortization formula
    monthly.payment <- ( principal * int.rate ) / 
                       (1 - (1 + int.rate)^(-months) )
    
    monthly.payment <- round( monthly.payment, 2 )
    
    return( monthly.payment )
 
 }
```

```{r ex="example-04", type="sample-code", tut=TRUE}
# Try these:

calcMortgage( 100000, 30, 0.05 )

calcMortgage( 30, 100000, 0.05 )
```
